---
title: "Lab 10"
author: "Vanessa Tang"
date: "3/11/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 10: R Tutorial on Positions and Roles in Networks

## 1. Reading in Example Network Data

```{r}
library(igraph)

class182_networkdata=read.csv(file="https://raw.githubusercontent.com/JeffreyAlanSmith/Integrated_Network_Science/master/data/class182_networkdata.csv")
head(class182_networkdata)
```

```{r}
class182_attributes=read.csv(file="https://raw.githubusercontent.com/JeffreyAlanSmith/Integrated_Network_Science/master/data/class182_attributedata.csv")
head(class182_attributes)
```

Friendship
```{r}
edgelist_friendship=class182_networkdata[class182_networkdata$friend_tie>0, 
                                         c("ego", "alter", "friend_tie")]
head(edgelist_friendship)
```

```{r}
net182_friend=graph_from_data_frame(d=edgelist_friendship, directed=T, 
                                    vertices=class182_attributes) 
net182_friend
```

Social Interaction Network
```{r}
edgelist_social=class182_networkdata[class182_networkdata$social_tie>0, 
                                     c("ego", "alter", "social_tie")]
```

Social Network
```{r}
net182_social=graph_from_data_frame(d=edgelist_social, directed=F, 
                                    vertices=class182_attributes) 

# remove repeat interactions
net182_social=simplify(net182_social, edge.attr.comb="mean") 

net182_social
```

Task Network
```{r}
edgelist_task=class182_networkdata[class182_networkdata$task_tie>0, 
                                   c("ego", "alter", "task_tie")]

net182_task=graph_from_data_frame(d=edgelist_task, directed=F, 
                                  vertices=class182_attributes) 

# again remove repeats
net182_task=simplify(net182_task, edge.attr.comb="mean")

net182_task
```

## 2. Getting the Data Ready

Matrices for each relation: friend, social, task
- Helps define structural equivalence
```{r}
friend_mat=as_adjacency_matrix(net182_friend, attr="friend_tie", sparse=F) 
social_mat=as_adjacency_matrix(net182_social, attr="social_tie", sparse=F)
task_mat=as_adjacency_matrix(net182_task, attr="task_tie", sparse=F)

friend_mat
```

Matrix of incoming ties = transpose friend matrix
```{r}
friend_mat_in=t(friend_mat)
```

Combine info
- Standardize columns for comparison: divide by st dev
- Mean center

```{r}
# subtract mean and divide by st dev
friend_mat_std=(friend_mat-mean(friend_mat))/sd(friend_mat) #for friendship
friend_mat_in_std=t(friend_mat_std) #for the incoming friendship ties
task_mat_std=(task_mat-mean(task_mat))/sd(task_mat) #for task
social_mat_std=(social_mat-mean(social_mat))/sd(social_mat) #for social

# combine
friends_task_social_std=cbind(friend_mat_std, friend_mat_in_std, 
                              task_mat_std, social_mat_std) 
dim(friends_task_social_std) 
```

```{r}
friends_task_social=cbind(friend_mat, friend_mat_in, task_mat, social_mat) 
```

## 3. Calculating Distance Between Nodes in the Network based on Pattern of Ties

```{r}
# euclidean distance matrix: smaller means more similar ties
euclid_dist=dist(x=friends_task_social_std, method="euclidean") 
as.matrix(euclid_dist)[,1:5]
```

```{r}
fit=cmdscale(euclid_dist, k=2) 

x=fit[,1]
y=fit[,2]

plot(x, y, xlab="Coordinate 1", ylab="Coordinate 2", 
     main="2 Dimensional MDS Solution", type="n") 
text(x, y, labels=1:length(x), cex=1)
```

## 4. Defining Positions

- Nodes are placed into positions based on having the same pattern of ties to other nodes
- Could this be helpful for Capstone? Students that have taken similar classes are places in similar positions

### 4.1. Hierarchical Clustering

```{r}
hc=hclust(euclid_dist)
plot(hc)
```

```{r}
plot(hc)
plot.clusters=rect.hclust(hc, h=12)
```

```{r}
plot(hc)
plot.clusters=rect.hclust(hc, h=8)
```

- the number of clusters increases as we lower the distance threshold

```{r}
hc_ids=cutree(hc, h=12)  
hc_ids
```

```{r}
library(NbClust)

clusters=NbClust(data=friends_task_social_std, distance="euclidean",
                 method="complete", index=c("ch"))  
clusters 
```
- how is the "optimal solution" defined?

## 5. Role Analysis

### 5.1 Permuting the Matrix by Position

```{r}
id_dat=data.frame(ids=class182_attributes$ids, position=hc_ids)
id_dat=id_dat[order(id_dat$position),]

# Using matrices from above: social
social_mat_rearrange=social_mat[id_dat$ids, id_dat$ids]

library(RColorBrewer)

column_cols=c("red", "green", "black", "blue")

heatmap(social_mat_rearrange, Rowv=NA, Colv=NA, revC=T,
        col=colorRampPalette(brewer.pal(6, "Blues"))(25),
        ColSideColors=column_cols[id_dat$position],
        RowSideColors=column_cols[id_dat$position], symm=T)
```

- most social interactions happen within position (within same color)

### 5.2 Constructing a Blockmodel

- Blockmodel: a matrix where the positions themselves are the nodes, or blocks, of interest
- All actors in the ‘red’ position (1, 9, 10, 12 and 15 in the figure above) get collapsed into a single node, or block, labeled, “Block 1
- Shows density of within and between block ties

```{r}
detach(package:igraph)
library(sna)

blockmod_friend=blockmodel(friend_mat, ec=hc_ids)
blockmod_task=blockmodel(task_mat, ec=hc_ids)
blockmod_social=blockmodel(social_mat, ec=hc_ids)

blockmod_friend
```

### 5.3 Plotting a Blockmodel

```{r}
library(ggraph)
library(tidygraph)

block_model_edgelist_function=function(block_model, relation_label, 
                                       directed=T){ 
#Inputs are the blockmodel, the name of the relation, 
#and if the relation is assumed to be undirected

#First we grab the actual matrix from the blockmodel
block_mat=block_model$block.model

#Here we set any NA to 0
block_mat[is.na(block_mat)]=0

#Now we create a little network based on the matrix
net_block=network(block_mat, loops=T, directed=directed)

#Here we extract the edgelist:
edges_netblock=as.edgelist(net_block)

#Now we get the edge weights, stringing out the matrix
#into a vector. We only extract those weights corresponding
#to where an edge exists, defined by the edgelist 
#extracted above.
weight_edge=c(block_mat[edges_netblock])

#Now we create a little data frame putting the information together.
block_edgelist=data.frame(edges_netblock, weight=weight_edge, 
                          Tie=relation_label)

#Here we create the additional weighting scheme,
#where weight is equal to 0, 1, or 2, depending if it is
#less than the mean, greater than the mean (but less than 1 sd above mean)
#or greater than 1 sd above the mean. 
edge_mean=mean(block_mat)
edge_sd=sd(block_mat)
edge_max=max(block_mat)

block_edgelist$WeightRecode=cut(block_edgelist$weight, 
    breaks=c(0, edge_mean, edge_mean+edge_sd, edge_max),
    include.lowest=T, right=F, labels=c("0", "1", "2")) 

colnames(block_edgelist)[1:2]=c("sender", "receiver")

block_edgelist
}

# create edgelists
blockedges_friend=block_model_edgelist_function(block_model=blockmod_friend, 
                                                relation_label="friendship", 
                                                directed=T)
blockedges_social=block_model_edgelist_function(block_model=blockmod_social, 
                                                relation_label="social", 
                                                directed=F)

blockedges_task=block_model_edgelist_function(block_model=blockmod_task, 
                                              relation_label="task", 
                                              directed=F)

blockedges_friend
```

Simple plot
```{r}
block_dat=rbind(blockedges_friend, blockedges_social, blockedges_task)
block_dat=block_dat[block_dat$WeightRecode %in% c("1", "2"),]
plot_net=tbl_graph(nodes=data.frame(ids=paste("Block", 1:4)), 
                   edges=block_dat, directed=F)

ggraph(plot_net, layout='circle') + 
  geom_edge_parallel(aes(colour=Tie))+
  geom_node_text(aes(label=ids), 
                 nudge_x=c(.15, 0, -.15, 0),                         
                 nudge_y=c(0, .10, 0, -.10))+ 
  theme_graph()
```

Plot with loops
```{r}
ggraph(plot_net, layout='circle') + 
  geom_edge_parallel(aes(colour=Tie))+
  geom_edge_loop(aes(colour="friendship", span=45, strength=.5))+
  geom_edge_loop(aes(colour="social", direction=90, span=45, strength=.5))+
  geom_edge_loop(aes(colour="task", direction=135, span=45, strength=.5))+
  geom_node_text(aes(label=ids), 
                 nudge_x=c(.16, .17, -.16, 0),                         
                 nudge_y=c(-.03, -.02, -.03, -.10))+
  theme_graph()
```

Plot with loops and weights
```{r}
ggraph(plot_net, layout='circle') + 
  geom_edge_parallel(aes(colour=Tie, width=WeightRecode), 
                     sep=unit(2, "mm"))+
  geom_edge_loop(aes(colour="friendship", span=45, 
                     strength=.5, width=WeightRecode))+
  geom_edge_loop(aes(colour="social", direction=90, span=45, 
                     strength=.5, width=WeightRecode))+
  geom_edge_loop(aes(colour="task", direction=135, span=45, 
                     strength=.5, width=WeightRecode))+
  scale_edge_width_manual(values=c(1, 2))+
  geom_node_text(aes(label=ids), 
                 nudge_x=c(.17, .18, -.17, 0),                         
                 nudge_y=c(-.05, -.05, -.05, -.10))+
  theme_graph()
```

### 5.4 Interpretation of Blockmodel

Block 4
- no friendship ties
- socializes with block 1 and block 2 
- heavy task interactions with block 1 and block 2
- do more tasks than others in the classroom: calculate total number of tasks for each block

Total level of task interactions for each block:
```{r}
rowSums(blockmod_task$block.model, na.rm=T)
```

Block 3
- mostly friends and socialize with others in block 3
- much lower levels of social interaction and task interactions with block 4, the teacher (compared to the level of contact between block 4 and block 1 and 2)

Task interactions with block 4
```{r}
round(blockmod_task$block.model[,4], 3)
```

Gender and race for block 1 (high engagement)
- high social interactions with each other 
- very high levels of task engagement with the teacher
```{r}
table(class182_attributes$gender[hc_ids==1])
table(class182_attributes$race[hc_ids==1])
```

Gender and race for block 3 (low engagement)
- high internal social interaction 
- low task engagement with the teacher
```{r}
table(class182_attributes$gender[hc_ids==3])
table(class182_attributes$race[hc_ids==3])
```

# 6. Local Equivalence

Triad census
```{r}
source("https://www.uni-due.de/hummell/rolecensus/f.rc.r")

task_mat_binary=task_mat
task_mat_binary[task_mat_binary>0]=1

task_triads_ego=f.rc(task_mat_binary)

dim(task_triads_ego)
task_triads_ego[1:10, 1:10]
```

Use triad census matrix to see which nodes have similar patterns of task interactions
- concor: clusters nodes into positions based on the correlations between rows

```{r}
source("https://raw.githubusercontent.com/JeffreyAlanSmith/Integrated_Network_Science/master/R/concor_functions.R")

source("https://raw.githubusercontent.com/JeffreyAlanSmith/Integrated_Network_Science/master/R/clustConfigurations_concor.R")

task_triads_ego_transpose=t(task_triads_ego)
concor_results=concor(task_triads_ego_transpose, depth=6)
concor_results
```

Calculate the observed correlation matrix and then produce the plot of fit statistics
```{r}
observed_correlation=cor(task_triads_ego_transpose)
diag(observed_correlation)=NA

concor_fit=clustConfigurations_concor(concor_results, observed_correlation) 
```

```{r}
positions_triads=concor_results[,4]
positions_triads
```

```{r}
position_triad_mat=aggregate(task_triads_ego, 
                             by=list(cluster=positions_triads), FUN=mean)
position_triad_mat[,c(1, 2, 7, 8, 17, 18)]
```
