---
title: "Lab 8"
author: "Vanessa Tang"
date: "2/21/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 8: R Tutorial on Network Cohesion and Communities

## 1. Setting up the Session and Reading in First Example Network

```{r}
library(igraph)
library(ggplot2)

class182_networkdata=read.csv(file="https://raw.githubusercontent.com/JeffreyAlanSmith/Integrated_Network_Science/master/data/class182_networkdata.csv")
head(class182_networkdata) 

class182_attributes=read.csv(file="https://raw.githubusercontent.com/JeffreyAlanSmith/Integrated_Network_Science/master/data/class182_attributedata.csv")
class182_attributes

edgelist_friendship=class182_networkdata[class182_networkdata$friend_tie>0,
                                         c("ego", "alter", "friend_tie")]
head(edgelist_friendship)

net182_friend=graph_from_data_frame(d=edgelist_friendship, directed=T,
                                      vertices=class182_attributes) 
net182_friend 
```

## 2. Cohesion

Plot
```{r}
plot(net182_friend, vertex.label=NA, edge.arrow.size=.5, 
     edge.arrow.width=.5, edge.color="light gray", 
     edge.curved=.2, vertex.frame.color= NA, margin=0)
```

```{r}
# density
edge_density(net182_friend)

# components
components_friendship=components(graph=net182_friend, mode="weak")
components_friendship

# proportion of nodes in each component
components_friendship$csize/sum(components_friendship$csize) 

# biconnected components
bicomponents_friendship=biconnected_components(graph=net182_friend)
bicomponents_friendship$components
```

```{r}
net182_friend_und=as.undirected(net182_friend, mode='collapse') 

net182_friend_und
```

Vertex disjoint: node to remove to disconnect source and target
```{r}
vertex_disjoint_paths(graph=net182_friend_und, source=1, target=9)
```

Vertex connectivity: num nodes to remove to disconnect
```{r}
vertex_connectivity(graph=net182_friend_und)

# remove isolates and recalc
isolates=which(degree(net182_friend_und)==0)

net182_noisolates=delete_vertices(net182_friend_und, isolates)

vertex_connectivity(graph=net182_noisolates)
```

## 3. Community (or Group) Detection

### 3.1. Walktrap

4 steps
```{r}
friend_comm_wt4=cluster_walktrap(graph=net182_friend_und, steps=4, 
                                 membership=T)

mems_wt_4step=membership(friend_comm_wt4)
mems_wt_4step

mod_wt_4step=modularity(friend_comm_wt4)
mod_wt_4step
```

3 steps
```{r}
friend_comm_wt3=cluster_walktrap(graph=net182_friend_und, steps=3, 
                                 membership=T)
mems_wt_3step=membership(friend_comm_wt3)
mems_wt_3step

mod_wt_3step=modularity(friend_comm_wt3)
mod_wt_3step
```

Compare
```{r}
table(mems_wt_4step, mems_wt_3step)
```

```{r}
par(mfrow=c(1, 2))
layout=layout.fruchterman.reingold(net182_friend) 

plot(net182_friend_und, layout=layout, #note the use of layout
     vertex.color=mems_wt_4step, edge.arrow.size=.5,
     edge.arrow.width=.5, edge.color="light gray", 
     vertex.size=20, main="Walk Trap: 4 Steps", margin=0)

plot(net182_friend_und, layout=layout, #note the use of layout
     vertex.color=mems_wt_3step, edge.arrow.size=.5, 
     edge.arrow.width=.5, edge.color="light gray", 
     vertex.size=20, main="Walk Trap: 3 Steps", margin=0)
```

### 3.2. Edge Betweenness

```{r}
friend_comm_eb=cluster_edge_betweenness(graph=net182_friend_und)
mems_eb=membership(friend_comm_eb)
table(mems_wt_3step, mems_eb) 
plot(as.dendrogram(friend_comm_eb))
```

```{r}
extract_modularity_data=function(communities, graph){

mems_list=list() #list where membership information will be saved
num_communities=NA #vector where number of communities will be saved
modularities=NA #vector to store modularity scores

#Extracting levels of aggregation, or steps,
#in the hierarchical merge data.
num_merges=0:nrow(communities$merges)

#Note that we start from 0 as the first level
#corresponds to all nodes in their own community,
#and that does not have a row in the  merge data frame.

#Looping through each level of aggregation

for (x in 1:length(num_merges)){
  
  #We first extract the membership 
  #information at the given merge level using a 
  #cut_at function. The inputs are the community
  #object and the merge step of interest. 
  
  mems_list[[x]]=cut_at(communities, steps=num_merges[x])

  #Now we calculate the number of communities associated
  #with the given clustering solution: 
  num_communities[x]=length(unique(mems_list[[x]]))

  #Let's also calculate the modularity score, just to make sure
  #we get the right value for that set of community memberships:
  modularities[x]=modularity(graph, mems_list[[x]])
    }

#We will now put together our extracted 
#information in a data frame. 

plot_data=data.frame(modularity=modularities, 
                     num_communities=num_communities)

#Let's reorder to go from low number of communities to high:
mems_list=mems_list[order(plot_data$num_communities)]
plot_data=plot_data[order(plot_data$num_communities),] 
rownames(plot_data)=1:nrow(plot_data)

#outputting resuts in a list:
list(summary_data=plot_data, membership_list=mems_list)
}

modularity_data=extract_modularity_data(communities=friend_comm_eb, 
                                        graph=net182_friend_und)

summary_data=modularity_data[[1]]
summary_data
```

```{r}
ggplot(summary_data, aes(num_communities, modularity))+
  geom_bar(stat="identity", fill="grey")+
  geom_line(color="black", linetype="solid")+ 
  geom_point(shape=19, size=1, color="black")+
  labs(title="Modularity by Number of Communities",
       x="Number of Communities", y="Modularity")
```

```{r}
mems_list=modularity_data[[2]]
mems_ids=which(summary_data$num_communities==4) 
mems_eb4=mems_list[[mems_ids]]
mems_eb4
```

### 3.3. Multi-level Clustering

```{r}
friend_comm_multi=cluster_louvain(graph=net182_friend_und)
friend_comm_multi$memberships[,1:10]

mems_mult_level1=friend_comm_multi$memberships[1,]
mems_mult_level2=friend_comm_multi$memberships[2,]

table(mems_mult_level1, mems_mult_level2) 

friend_comm_multi$modularity
```

```{r}
par(mfrow=c(1,2))

#4 communities
plot(net182_friend_und, layout=layout, vertex.color=mems_mult_level2,
     edge.arrow.size=.5,edge.arrow.width=.5, 
     edge.color="light gray", vertex.size=20,
     main="Level 2", margin=0)

#5 communities
plot(net182_friend_und, layout=layout, vertex.color=mems_mult_level1, 
     edge.arrow.size=.5,edge.arrow.width=.5, 
     edge.color="light gray", vertex.size=20,
     main="Level 1", margin=0)
```

### 3.4. Cohesive Blocking

```{r}
friend_comm_cohesive=cohesive_blocks(graph=net182_friend_und)
friend_comm_cohesive
```

```{r}
plot(friend_comm_cohesive, net182_friend_und) 
```

## 4. Demographic Characteristics

```{r}
library(car)
cols=recode(class182_attributes$race, as.factor=F, "c('black')='blue'; 
            else='white'")
table(class182_attributes$race,cols) 
```

```{r}
plot(friend_comm_eb, net182_friend_und, col=cols, layout=layout, 
     main="Race Mapped onto Communities", margin=0)
```

```{r}
proportion_function=function(communities, attribute){ #inputs are community 
#membership and the attribute vector
#Here we calculate the proportion that fall into each category 
#found in attribute. We first do a table and then find the proportion 
#in each category. This is done for each community (using the tapply 
#option over the communities).

dat=tapply(factor(attribute), communities, function(x) {y=table(x);y/sum(y)})

do.call(rbind, dat) #We then output it as a matrix using a do.call 
#command and rbind option
} 

proportion_function(communities=mems_eb, attribute=class182_attributes$race)
```

```{r}
table_list=list()
for (i in 1:max(mems_eb)){
tab=table(class182_attributes$race[mems_eb==i])
table_list[[i]]=tab/sum(tab)
}

do.call(rbind, table_list)

table(class182_attributes$race)
```

## 5. Community Overlap

```{r}
class182_dyads=class182_networkdata[,c("ego", "alter")]
head(class182_dyads)

mat182_friend=as_adjacency_matrix(net182_friend_und, sparse=F)
diag(mat182_friend)=NA 
friend=c(t(mat182_friend))  

same_comm=mems_mult_level2[class182_dyads$ego]==mems_mult_level2[class182_dyads$alter]  

table(same_comm, friend)
```

Odds ratio: odds of 2 people in same community choosing each other as friends
```{r}
(70*134)/(14*22)
```

## 6. Big Data Example (Supplementary Section)

```{r}
email_edges=read.table(file="https://raw.githubusercontent.com/JeffreyAlanSmith/Integrated_Network_Science/master/data/email-EuAll.txt")

dim(email_edges)
colnames(email_edges)=c("ego", "alter")
email_edges=email_edges+1
email_edges=email_edges[email_edges[,1]!=email_edges[,2],]
email_network=graph_from_data_frame(d=email_edges, directed=T, 
                                    vertices=(id=as.numeric(1:265214)))
email_network=as.undirected(email_network, mode="collapse")
```
- error reading in data

```{r}
email_network=as.undirected(email_network, mode="collapse")
email_mems=email_comm_multi$memberships
nrow(email_mems) 

length.func=function(x){length(unique(x))} 
num_comms=apply(email_mems, 1, length.func) 
num_comms 

email_comm_multi$modularity
```

```{r}
mems_email_level4=email_comm_multi$memberships[4,]
summary(as.numeric(table(mems_email_level4))) 
```

```{r}
subcommunity_function=function(graph, initial_communities, community_number){
#The inputs will be the original network, 
#the original community/group memberships and 
#the community/group number of interest (i.e., the community 
#that you want to divide further).

#here will create a subgraph of just the community of interest
subgraph1=induced_subgraph(graph=graph, vids=which(initial_communities==community_number)) 

#We now perform a community detection algorithm (using fast and greedy) 
#on the subgraph
comm1=cluster_fast_greedy(graph=subgraph1) 

mems_subgraph1=membership(comm1) #grabbing the community membership of each 
#person in the subgraph

#Now we grab the ids of those in the subgraph, so we can map them back 
#onto the original, full network 
ids_map=as.numeric(vertex_attr(subgraph1,"name")) 

mems_new=initial_communities #just copying the original communities

#Here, we begin to relabel the communities so we can put them back onto the 
#original set of communties on the full network. We want to make sure that 
#these new community ids are unique, so we take the max
#original community number and add that to the community ids on the subgraph.
mems_subgraph1_relabel=mems_subgraph1+max(initial_communities) 

#Here we put the new communities onto a vector of community 
#membership corresponding to the whole network. 
mems_new[ids_map]=mems_subgraph1_relabel 

#Note we just change those in the subgraph of interest.

#We can then relabel all communities, if desired, to take out the old 
#community number and put in order from low to high:
num_comms_new=length(unique(mems_new))
mems_updated=as.numeric(as.character(factor(mems_new, 
                                            labels=1:num_comms_new)))

#here we output the subgraph, the membership and the updated
#vector of community membership:
list(subgraph=subgraph1,mems_subgraph=mems_subgraph1, 
     membership_updated=mems_updated)
}

subcommunity_dat=subcommunity_function(graph=email_network, 
                        initial_communities=mems_email_level4,
                        community_number=2655) 

subnet=subcommunity_dat$subgraph
mems_subnet=subcommunity_dat$mems_subgraph
mems_updated=subcommunity_dat$membership_updated

layout_email=layout.fruchterman.reingold(subnet)

plot(subnet, vertex.label=NA, vertex.size=.6, layout=layout_email, 
     edge.arrow.size=.5, edge.arrow.width=.5, edge.color="light gray", 
     edge.curved=.2, vertex.frame.color=NA, 
     vertex.color=mems_subnet, margin=0)
```

```{r}
data.frame(table(mems_updated[mems_email_level4==2655]))
modularity(email_network, mems_updated)

subcommunity_dat=subcommunity_function(graph=email_network,
                                      initial_communities=mems_updated,
                                      community_number=33)  
```
