---
title: "Lab 9"
author: "Vanessa Tang"
date: "2/21/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 9: R Tutorial on Network Centrality and Hierarchy

## 1. Setting up the Session and Reading in Example Network Data

```{r}
library(igraph)

social_data=read.csv(file="https://raw.githubusercontent.com/JeffreyAlanSmith/Integrated_Network_Science/master/data/social_interactions_s641.csv")
head(social_data) 

edgelist_social=social_data[social_data$social_tie>0, ] 
head(edgelist_social)

# create igraph object
s641_social=graph_from_data_frame(d=edgelist_social, directed=T, 
                                    vertices=(id=1:22)) 

# task data
task_data=read.csv(file="https://raw.githubusercontent.com/JeffreyAlanSmith/Integrated_Network_Science/master/data/task_interactions_s641.csv")
head(task_data)

edgelist_task=task_data[task_data$task_tie>0, ] 

s641_task=graph_from_data_frame(d=edgelist_task, directed=T, 
                                  vertices=(id=1:22)) 
```

Plotting networks
```{r}
par(mfrow=c(1,2)) 

plot(s641_social, vertex.frame.color=NA, edge.arrow.size=.25, 
    vertex.size=8, main="Social Interactions", margin=0)

plot(s641_task, vertex.frame.color=NA, edge.arrow.size=.25, 
     vertex.size=8, main="Task Interactions", margin=0)
```
- Really cool to compare and see differences between social vs task

## 2. Centrality

### 2.1 Centrality Measures for Social Interactions

Indegree
```{r}
indegree_social=degree(graph=s641_social, mode="in")
data.frame(indegree_social)
```

Outdegree
```{r}
outdegree_social=degree(graph=s641_social, mode="out")
data.frame(outdegree_social)
```

Out closeness: mean distance of each node to all other nodes
```{r}
outcloseness_social=closeness(graph=s641_social, mode="out", normalized=F)
data.frame(outcloseness_social)
```

In closness
```{r}
incloseness_social=closeness(graph=s641_social, mode="in", normalized=F)
data.frame(incloseness_social)
```

Can use distance matrix to solve problem of infinite distances (isolates)
```{r}
dist_mat_social=distances(graph=s641_social, mode="out") #calculating the 
#distance matrix
diag(dist_mat_social)=NA #ignoring the diagonal
dist_mat_social_inverted=1/dist_mat_social #getting inverted distance matrix

outcloseness_social2=apply(dist_mat_social_inverted, MARGIN=1, 
                           FUN=mean, na.rm=T) 
data.frame(outcloseness_social2)
```

Comparing different ways of calculating out closeness (function vs distance matrix)
```{r}
cor(outcloseness_social, outcloseness_social2) 
```

Another way to compute in closness using distance matrix
```{r}
incloseness_social2=apply(dist_mat_social_inverted, MARGIN=2,
                          FUN=mean, na.rm=T) 
data.frame(incloseness_social2)
```

Betweenness centrality: num of shortest paths going through vertex
```{r}
betweenness_social=betweenness(graph=s641_social, normalized=F) 
data.frame(betweenness_social)
```

Eigenvector centrality: greater weight to node when more connected to other highly connected nodes
- measuring node network importance
```{r}
s641_social_undirected=as.undirected(s641_social, mode="collapse") 

# eigenvector centrality
ev_obj_social=eigen_centrality(s641_social_undirected)
eigen_social=ev_obj_social$vector

data.frame(eigen_social)
```

```{r}
ids=V(s641_social)$name

central_social=data.frame(ids=ids, net="social",
                            indegree=indegree_social,
                            outdegree=outdegree_social,
                            incloseness=incloseness_social, 
                            incloseness2=incloseness_social2, 
                            outcloseness2=outcloseness_social2,
                            between=betweenness_social, 
                            eigen=eigen_social)

apply(central_social[, -c(1,2)], MARGIN=2, FUN=order, decreasing=T)
```

```{r}
library(reshape)

social_long=melt(central_social[,c("ids", "indegree", "incloseness2",
                                   "between", "eigen")], id.vars="ids")  
head(social_long)

library(ggplot2)
ggplot(social_long, aes(x=factor(ids, levels=1:length(ids)), y=value))+
  geom_bar(stat="identity", position="dodge")+
  facet_wrap(~variable, scales="free_y", ncol=2)+
  xlab("Ids")+ylab("Centrality")+
 theme(axis.text=element_text(size=6.5))
```

```{r}
plot(s641_social, vertex.size=central_social$between/5, 
     vertex.label=V(s641_social)$name, 
     edge.arrow.size=0.25,
     layout=layout.fruchterman.reingold, 
     main="Classroom S641 Social Talk", margin=0) 
```

### 2.2. Correlations between Centrality Measures

```{r}
cor_tab1=cor(central_social[,-c(1,2)])
round(cor_tab1[,c("indegree", "outdegree", "incloseness2", 
                            "outcloseness2", "between", "eigen")], 3)
```

### 2.3. Centrality for Task Interactions

```{r}
indegree_task=degree(s641_task, mode="in")
outdegree_task=degree(s641_task, mode="out")
incloseness_task=closeness(s641_task, mode="in", normalized=F)
outcloseness_task=closeness(s641_task, mode="out", normalized=F)

dist_mat_task=distances(graph=s641_task, mode="out")
diag(dist_mat_task)=NA
dist_mat_task_inverted= 1/dist_mat_task

outcloseness_task2=apply(dist_mat_task_inverted, MARGIN=1, FUN=mean, na.rm=T)
incloseness_task2=apply(dist_mat_task_inverted, MARGIN=2, FUN=mean, na.rm=T) 

betweenness_task=betweenness(s641_task, normalized=F)

s641_task_undirected=as.undirected(s641_task, mode="collapse")
ev_obj_task=evcent(s641_task_undirected)
eigen_task=ev_obj_task$vector

central_task=data.frame(ids=ids, net="task",
                          indegree=indegree_task,
                          outdegree=outdegree_task,
                          incloseness=incloseness_task,
                          incloseness2=incloseness_task2, 
                          outcloseness2=outcloseness_task2, 
                          between=betweenness_task, 
                          eigen=eigen_task)

apply(central_task[, -c(1,2)], MARGIN=2, FUN=order, decreasing=T)
```

```{r}
plot(s641_task, vertex.size=central_task$indegree, 
     vertex.label=V(s641_social)$name, edge.arrow.size=0.25,
     layout=layout.fruchterman.reingold, 
     main="Classroom S641 Task Interactions", margin=0)
```

### 2.4. Task/Social Correlations

```{r}
task_long=melt(central_task[,c("ids", "indegree", "incloseness2",
                               "between", "eigen")], id.vars="ids") 
head(task_long)

task_long$net="task"
social_long$net="social"

social_task_long=rbind(social_long, task_long)

ggplot(social_task_long, 
       aes(x=factor(ids, levels=1:length(ids)), y=value, fill=net))+
  geom_bar(stat="identity", position="dodge")+
  facet_wrap(~variable, scales="free_y", ncol=2)+ 
  scale_fill_discrete(name="Network", breaks=c("social", "task"), 
                      labels=c("Social", "Task"))+
  xlab("Ids")+ylab("Centrality")+
  theme(axis.text=element_text(size=6.5))

```

```{r}
cor_tab2=cor(central_social[,-c(1,2)], central_task[,-c(1,2)])
rownames(cor_tab2)=paste(rownames(cor_tab2), "social", sep="_")
colnames(cor_tab2)=paste(colnames(cor_tab2), "task", sep="_") 
round(cor_tab2[,c("eigen_task", "indegree_task", "outdegree_task")], 3)
```

## 3. Centralization and the Distribution of Centrality

```{r}
summary(indegree_social)
summary(indegree_task)
sd(indegree_social)
sd(indegree_task)
```

```{r}
cent_social=centr_degree(graph=s641_social, mode="in", 
                         loops=FALSE, normalized=TRUE)
cent_social
```

```{r}
sum(max(indegree_social)-indegree_social)/sum(21-rep(0,21)) 
```

```{r}
cent_task=centr_degree(graph=s641_task, mode="in", loops=FALSE, 
                       normalized = TRUE)
cent_task
```

```{r}
den_social=density(indegree_social, from=0) #from is set to 0 as indegree
#cannot be less than 0.
den_task=density(indegree_task, from=0)

# plot
plot(range(den_social$x, den_task$x), range(den_social$y,den_task$y), 
     type="n", xlab="degree", ylab="density", 
     main="Indegree Distribution for Social and Task Networks")
lines(den_social, col="red" , lty=2, lwd=2)
lines(den_task, col="light blue", lty=2, lwd=2)
legend("topright", c("Social", "Task"), col=c("red", "light blue"), 
       lty=2, lwd=2) 
```
- task network much more centralized

## 4. Clustering and Hierarchy: Tau Statistic

```{r}
library(intergraph)

s641_social_network=asNetwork(s641_social)

source(file="https://raw.githubusercontent.com/JeffreyAlanSmith/Integrated_Network_Science/master/R/tau_functions.R")

library(ergm)
library(sna)
```

```{r}
weight_vector_rankedcluster=c(1, 0, 1, 1, 1, 0, 0, 0, 1, 0, 0, 1, 1, 0, 0, 1)
weight_vector_cluster=c(1, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
weight_vector_balance=c(0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)

tau_rankedcluster=tau_stat_function(network=s641_social_network, 
                                    weight.vector=weight_vector_rankedcluster)
tau_rankedcluster 
```

```{r}
tau_cluster=tau_stat_function(network=s641_social_network, 
                              weight.vector=weight_vector_cluster)
tau_cluster
```

```{r}
tau_balance=tau_stat_function(network=s641_social_network, 
                              weight.vector=weight_vector_balance)
tau_balance
```

```{r}
tau_rankedcluster[[2]][rownames(tau_rankedcluster[[2]]) %in% 
                         c("triadcensus.021D", "triadcensus.021U", 
                           "triadcensus.030T", "triadcensus.120D",  
                           "triadcensus.120U"), ]
```
