---
title: "Labs 4 and 5"
author: "Vanessa Tang"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 4: R Tutorial on Missing Network Data

### Part 1: Network Data with Missing Nodes

```{r}
# missing nodes
sorority_attributes=read.csv(file="https://raw.githubusercontent.com/JeffreyAlanSmith/Integrated_Network_Science/master/data/sorority_attributes_wmissingdata.csv")

# friendships
sorority_matrix=read.csv(file="https://raw.githubusercontent.com/JeffreyAlanSmith/Integrated_Network_Science/master/data/sorority_matrix_wmissingdata.csv", row.names=1)
sorority_matrix=as.matrix(sorority_matrix)
```

Look at missing data: `sorority_attributes$missing` == 1
```{r}
missing=which(sorority_attributes[,"missing"]==1)
missing
```

Look at not missing data: `sorority_attributes$missing` == 0
```{r}
notmissing=which(sorority_attributes[,"missing"]==0)
notmissing
```

First missing node: 2
- Full of NA because node 2 is missing (no info)
```{r}
sorority_matrix[2, 1:15]
```

Look at column for missing node 2: show if other people have info for node 2
- some info from other nodes: 0 no relationship, 1 relationship, NA missing node
- don't know if node 2 would have also nominated nodes but know if another nonmissing node nominated node 2
```{r}
print(sorority_matrix[1:15, 2])
print(sorority_matrix[2, 7])
```

### Part 2: Constructing the Network Using no Imputation

Remove missing nodes
```{r}
sorority_matrix_removemissing=sorority_matrix[-missing, -missing]
dim(sorority_matrix_removemissing)
```

Create igraph object
```{r}
library(igraph)

sororitynet_noimpute=graph_from_adjacency_matrix(
  adjmatrix=sorority_matrix_removemissing, mode="directed")

sororitynet_noimpute
```

Plot
```{r}
plot(sororitynet_noimpute, vertex.label=NA, vertex.size=10, 
     edge.arrow.size=.5, edge.arrow.width=.5)
```

Mean closeness
1. Define shortest path
2. Take inverse
3. Mean
```{r}
# shortest path
dist_noimpute=distances(graph=sororitynet_noimpute, mode="out")
diag(dist_noimpute)=NA 

# inverse and mean
mean_closeness_noimpute=mean(1/dist_noimpute, na.rm=T) 
mean_closeness_noimpute
```

Creating closness function: same as above but easier
```{r}
closeness_function=function(network){
dist_matrix=distances(graph=network, mode="out")
diag(dist_matrix)=NA
mean_closeness=mean(1/dist_matrix, na.rm=T)
mean_closeness
}

closeness_function(sororitynet_noimpute)
```

### Part 3: Gauging the Level of Bias

#### 3.1 Gauging Bias Using Past Results
Expected bias = .25
Table 6 (in Smith and Moody 2013)

#### 3.2 Gauging Bias Using Predictive Models
http://www.soc.duke.edu/~jmoody77/missingdata/calculator.htm
- size = 72 (num nodes)
- percent missing = 14/72 = 20%
- indegree standard deviation = 1.61 (see below)
- directed/undirected = directed
- type of missing data
- correlation between missing nodes and centrality at .25

Expected bias = .19
```{r}
indeg=degree(sororitynet_noimpute, mode="in")
sd(indeg)
```

### Part 4: Simple Imputation Options: Network Reconstruction

Missing nodes with nominations
```{r}
# nominations to missing node
nominations_missing_nodes=sorority_matrix[,missing]

# num nominations to each missing node
indeg_missing=colSums(nominations_missing_nodes, na.rm=T)

# find missing nodes that received at least 1 nomination
# only keep ones that got at least 1 nomination
impute_nodes=missing[indeg_missing>0]
impute_nodes
```

Missing nodes without nominations
```{r}
# nodes to remove: missing nodes with no nominations
still_missing=missing[indeg_missing==0]
still_missing
```

#### 4.1 Network Reconstruction: Directed (Asymmetric) Option

Assume missing nodes have no output ties, only use input ties to missing node

```{r}
sorority_matrix_impute_directed=sorority_matrix 

# outgoing ties from missing nodes = 0
sorority_matrix_impute_directed[missing,]=0 

# imputed matrix
sororitynet_impute_directed=graph_from_adjacency_matrix( 
  adjmatrix=sorority_matrix_impute_directed, mode="directed")
```

Plot
- red: missing
- blue: not missing

```{r}
cols=ifelse(sorority_attributes[,"missing"]==1, "red", "blue")
table(cols, sorority_attributes[,"missing"])

# add colors to network
V(sororitynet_impute_directed)$color=cols 

# remove nodes with no nominations (still_missing)
sororitynet_impute_directed=delete_vertices(graph=sororitynet_impute_directed, 
                                            v=still_missing)

plot(sororitynet_impute_directed, vertex.label=NA, vertex.size=7, 
     edge.arrow.size=.75, edge.arrow.width=.75)
```

Closeness
Closeness using directed imputation (.09) is lower than no imputation (.11)
```{r}
mean_closeness_impute_directed=closeness_function(sororitynet_impute_directed)
mean_closeness_impute_directed
```

#### 4.2 Network Reconstruction: Reciprocated Option

Assume all ties reciprocated: if there is an input tie to missing node, then missing node would also have output tie to receiver
- set outgoing ties from missing nodes equal to incoming ties to missing node: `nominations_missing_nodes`

```{r}
sorority_matrix_impute_recip=sorority_matrix 
head(nominations_missing_nodes)
```
- cols: missing nodes
- 1 = non-missing node (rows) nominated missing node (cols)

Need to transpose matrix to go from missing -> non-missing nodes
- set NA to 0: ties between missing nodes
- transpose to get missing to not missing ties
- set outgoing ties for missing nodes
```{r}
nominations_missing_nodes[is.na(nominations_missing_nodes)]=0
outgoing_ties_recip=t(nominations_missing_nodes) 
sorority_matrix_impute_recip[missing,]=outgoing_ties_recip

# look at node 2 (missing)
sorority_matrix_impute_recip[1:15, 2]         # outgoing ties
sorority_matrix_impute_recip[2, 13]           # new tie between 2 and 13
sorority_matrix_impute_directed[2, 13]        # originally no tie between 2 and 13
```

Create network
- remove still missing 
```{r}
sororitynet_impute_recip=graph_from_adjacency_matrix( 
  adjmatrix=sorority_matrix_impute_recip, mode="directed")
sororitynet_impute_recip=delete_vertices(graph=sororitynet_impute_recip, 
                                         v=still_missing)
sororitynet_impute_recip
```

Closeness
Mean closness using reciporcated ties imputation (.16) is higher than no imputation (.11) and directed imputation (.09)
```{r}
mean_closeness_impute_recip=closeness_function(sororitynet_impute_recip)
mean_closeness_impute_recip
```

#### 4.3 Network Reconstruction: Probabilistic Option

If tie exists from nonmissing to missing, assume tie exists from missing to nonmissing with probability p (reciprocity rate)

Reciprocity rate
- ratio: symmetric/(asymmetric + symmetric)
- 42% of the time when i nominates j, j nominates i
```{r}
p=reciprocity(graph=sororitynet_noimpute, mode="ratio")
p
```

Imputing
- transpose: nominations to missing nodes (rows = missing, cols = nonmissing)
- For each nomination from n (non-missing) to m (missing), we will take a draw from a binomial distribution with probability set to p
- Only draw when outgoing_ties_prob is equal to 1 (nonmissing node nominates missing node)
```{r}
sorority_matrix_impute_prob=sorority_matrix

# transpose
outgoing_ties_prob=t(nominations_missing_nodes)

set.seed(200)

# take draws from binomial
outgoing_ties_prob[outgoing_ties_prob==1]=
  rbinom(outgoing_ties_prob[outgoing_ties_prob==1], size=1, prob=p) 

# set outgoing ties for missing nodes
sorority_matrix_impute_prob[missing,]=outgoing_ties_prob
```

Creating network
- remove missing nodes with no incoming ties
```{r}
sororitynet_impute_prob=graph_from_adjacency_matrix( 
  adjmatrix=sorority_matrix_impute_prob, mode="directed")

sororitynet_impute_prob=delete_vertices(graph=sororitynet_impute_prob, 
                                        v=still_missing)
```

Closeness
Mean closeness using probabilistic imputing (.14) is higher than no imputation (.11) and simple directed imputation (.09) but lower than reciprocated imputation (.16)
```{r}
mean_closeness_impute_prob=closeness_function(sororitynet_impute_prob)
mean_closeness_impute_prob
```

#### 4.4 Comparing Estimates
```{r}
closeness_table=data.frame(true=.15, 
                           noimpute=mean_closeness_noimpute, 
                           direct=mean_closeness_impute_directed, 
                           recip=mean_closeness_impute_recip, 
                           prob=mean_closeness_impute_prob)

closeness_table
```

- reciprocated did best in this case
- directed was worse than doing nothing
- only know true because its simulated data