---
title: "Lab 5"
author: "Vanessa Tang"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Network Visualization

## Part 1: Basic Network Visualization

### 1.1: Setting up the Session

```{r}
# edges
class_edges=read.csv(file="https://raw.githubusercontent.com/JeffreyAlanSmith/Integrated_Network_Science/master/data/class555_edgelist.csv")

# attributes
class_attributes=read.csv(file="https://raw.githubusercontent.com/JeffreyAlanSmith/Integrated_Network_Science/master/data/class555_attributedata.csv")

library(igraph)

# create igraph object
class_net=graph_from_data_frame(d=class_edges, directed=T, 
                                vertices=class_attributes)
```

### 1.2: Basic Plotting of Network Data Using igraph

```{r}
# add color by gender
cols=ifelse(class_attributes$gender=="Female", "lightskyblue", "navy") 
V(class_net)$color=cols 
plot(class_net, margin=0)

# can also do it this way
# plot(class_net, vertex.color=cols)
```
Really helps to color code the nodes! Makes it much easier to see patterns: 2 groups of girls, 2 boys in girl groups, one main boy group.

Interested in number of ties to node (indegree)
Plot network:
- indegree as node size (+3) to make bigger and easier to see
- color vertex labels
```{r}
indeg=degree(class_net, mode="in")
plot(class_net, vertex.size=indeg+3, vertex.label.color="red", margin=-.10)
```

Making it prettier:
- remove vertex labels
- `vertex.frame.color`: remove black outline around nodes
- decrease arrow size
- make edges lighter grey: similar concept to data viz
- `edge.curved`: add curves to edges because it looks nicer
- change layout

```{r}
plot(class_net, vertex.size=indeg+3,vertex.label=NA,vertex.frame.color=NA,
     edge.arrow.size=.5, edge.arrow.width=.75, edge.color="light gray", 
     edge.curved=.15, layout=layout_with_kk, margin=-.10)
```

### 1.3: Plotting Network Data in R using Packages based on ggplot

ggplot: network object

```{r}
detach(package:igraph)
library(sna)
library(network)
library(ggplot2)
library(GGally)
library(intergraph)

# use intergraph to tranform igraph to network
class_net_sna=asNetwork(class_net)

set.vertex.attribute(class_net_sna, attrname="indeg", value=indeg)
```

#### 1.3.1: GGally Package

- `node.size`: by indegree
- `node.color`: by gender
- `edge.size`: make grey lines smaller
- `arrow.size`: make edge arrows smaller
- make edges lighter and smaller as before
- remove legend by using `+guides(size=FALSE)`
- `palette`: set node colors by m/f, automatically adds legend
- `edge.color` to c(“color”,“grey80”): will use node colors to color edges if m/m or f/f tie and grey for m/f and f/m (cross-group) ties

```{r}
ggnet2(class_net_sna,  node.size=indeg,  node.color="gender",
       palette=c("Male"="navy", "Female"="lightskyblue"),
       edge.size=.5, arrow.size=3, arrow.gap=0.02, 
       edge.color=c("color", "grey80"))+
  guides(size=FALSE)
```

#### 1.3.1: ggnetwork Package

Allows more control 
Syntax is very similar to normal ggplot plotting

```{r}
library(ggnetwork) 

ggplot(class_net_sna, arrow.gap=.015, 
        aes(x=x, y=y, xend=xend, yend=yend)) + 
  geom_edges(color="lightgray", 
             arrow=arrow(length=unit(7.5, "pt"), type="closed")) +
  geom_nodes(color=cols, size=indeg+3) +
  theme_blank() 
```

### 1.4: Contour Plots

Contour plots:
- topographical representation
- good for large and dense network
- `geom_density_2d`: reflects density of different regions

```{r}
ggplot(class_net_sna, arrow.gap=.01,
         aes(x=x, y=y, xend=xend, yend=yend)) + 
  geom_edges(color="lightgray",
             arrow=arrow(length=unit(5, "pt"), type="closed")) +
  geom_nodes(color=cols) +
  theme_blank() + 
  geom_density_2d()
```

Would be interested to see what this looks like in a bigger network. Almost seems more confusing to me now.

### 1.5: Interactive Plots

```{r}
library(networkD3)

# networkD3 indexes nodes from 0 not 1
class_edges_zeroindex=class_edges[,c("sender", "receiver")]-1
class_attributes_zeroindex=data.frame(id=class_attributes$id-1, indeg=indeg,
                                      gender=class_attributes$gender)
```

`forceNetwork`
- `Links` = edgelist of interest
-	`Nodes` = attributes
-	`Source` = name of variable on edgelist denoting sender of tie
-	`Target` = name of variable on edgelist denoting receiver of tie
-	`Group` = based on an attribute or network-based group (gender)
-	`Nodesize` = name of variable on attribute file to size nodes by (indegree)
-	`NodeID` = name of variable on attribute file showing id/name of node

```{r}
forceNetwork(Links=class_edges_zeroindex, Nodes=class_attributes_zeroindex,
             Source="sender", Target="receiver", Group="gender",
             Nodesize="indeg", NodeID="id", 
             opacity=0.9, bounded=FALSE, opacityNoHover=.2)
```

(Where is the html file?)

## Part 2: Visualizations for Dynamic Network Data

- good for continuous time, streaming data
- time stamped data

### 2.1: Preparing the Data

`edge_spells`: classroom interactions
- `start_time`
- `end_time`
- `send_col`: who initiated interaction
- `receive_col`: who received interaction

`vertex_spells`: movement of nodes in and out of network

```{r}
# read in data
edge_spells=read.csv(file="https://raw.githubusercontent.com/JeffreyAlanSmith/Integrated_Network_Science/master/data/example_edge_spells.csv")
vertex_spells=read.csv(file="https://raw.githubusercontent.com/JeffreyAlanSmith/Integrated_Network_Science/master/data/example_vertex_spells.csv")
attributes_example2=read.table("https://raw.githubusercontent.com/JeffreyAlanSmith/Integrated_Network_Science/master/data/class_attributes.txt", header=T)

# networkDynamic object
library(networkDynamic)

net_dynamic_interactions=networkDynamic(vertex.spells=vertex_spells, 
                                        edge.spells=edge_spells)
```

### 2.2: Time Flattened Visualizations

Create 2 networks: first 20 and last 20 minutes
```{r}
extract_nets=get.networks(net_dynamic_interactions, start=0, 
                           time.increment=20)
extract_nets
```

```{r}
cols_example2=ifelse(attributes_example2$gnd==2, "lightskyblue","navy")

# set location so nodes show up in same place in both plots
locs=network.layout.fruchtermanreingold(extract_nets[[1]], layout.par=NULL)

par(mfrow=c(1,2))

plot(extract_nets[[1]], main="Talk to Network, 0 to 20 Minutes", 
     vertex.col=cols, vertex.cex=2, coord=locs)

plot(extract_nets[[2]], main="Talk to Network, 20 to 40 Minutes", 
     vertex.col=cols, vertex.cex=2, coord=locs)
```

### 2.3 Dynamic Network Movies

```{r}
library(ndtv)

slice.par=list(start=0, end=43, interval=1, aggregate.dur=0, rule="latest")
set.network.attribute(net_dynamic_interactions, 'slice.par', slice.par)

# make and save movie as html
render.d3movie(net_dynamic_interactions, displaylabels=FALSE, vertex.cex=1.5, 
               output.mode="HTML", filename="classroom_movie1.html")
```

Increase aggregate duration to see interactions occurring within the minute
```{r}
slice.par=list(start=0, end=43, interval=1,  aggregate.dur=1, rule="latest")

set.network.attribute(net_dynamic_interactions, 'slice.par', slice.par)

render.d3movie(net_dynamic_interactions, displaylabels=FALSE, vertex.cex=1.5, 
               output.mode="HTML", filename="classroom_movie2.html")
```

Color nodes by gender
```{r}
slice.par=list(start=0, end=43, interval=1,  aggregate.dur=1, rule="latest")

set.network.attribute(net_dynamic_interactions, 'slice.par', slice.par)

render.d3movie(net_dynamic_interactions, displaylabels=FALSE, 
               vertex.cex=1.5, vertex.col=cols,
               output.mode="HTML", filename="classroom_movie3.html")
```

Aggregate every 10 min (tie exists between i and j if i and j interacted in first 10 min): 0-10, 10-20, 20-30, etc.
```{r}
slice.par=list(start=0, end=43, interval=10, aggregate.dur=10, rule="latest")

set.network.attribute(net_dynamic_interactions, 'slice.par', slice.par)

render.d3movie(net_dynamic_interactions, displaylabels=FALSE, 
               vertex.cex=1.5, vertex.col=cols,
               output.mode="HTML", filename="classroom_movie4.html")
```

Aggregate every 10 min but at an interval of 1: 0-10, 1-11, 2-12, etc.
```{r}
slice.par=list(start=0, end=43, interval=1, aggregate.dur=10, rule="latest")

set.network.attribute(net_dynamic_interactions, 'slice.par', slice.par)

render.d3movie(net_dynamic_interactions, displaylabels=FALSE, 
               vertex.cex=1.5, vertex.col=cols,
               output.mode="HTML", filename="classroom_movie5.html")
```

These dynamic network visualizations are so cool!! It's interesting to see how groups form and change. One thing I think that may be misleading is how node location changes over time. It seems like humans are naturally wired to asssociate locations, but sometimes the node location changed even when its ties didn't. 


