---
title: "Lab 7"
author: "Vanessa Tang"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Tutorial on Dyads and Triads

## 1. Setting up the Session and Reading in Example Network Data

```{r}
library(sna)

advice_data_frame=read.csv(file="https://raw.githubusercontent.com/JeffreyAlanSmith/Integrated_Network_Science/master/data/krackhardt_advice.csv")

head(advice_data_frame)

# subset for where tie existes: edgelist
advice_edgelist=advice_data_frame[advice_data_frame$advice_tie==1,]

attributes=read.csv(file="https://raw.githubusercontent.com/JeffreyAlanSmith/Integrated_Network_Science/master/data/krackhardt_attributes.csv")

attributes
```

Add attributes to edgelist
```{r}
size= nrow(attributes)
attr(advice_edgelist, "n")=size

attribute_list=do.call(list, attributes)
krack_advice=network(x=advice_edgelist, directed=T, 
                     vertex.attr=attribute_list) 
```

Friendship data
```{r}
friendship_data_frame=read.csv(file="https://raw.githubusercontent.com/JeffreyAlanSmith/Integrated_Network_Science/master/data/krackhardt_friendship.csv")

head(friendship_data_frame)

# subset for where tie exists
friendship_edgelist=
  friendship_data_frame[friendship_data_frame$friendship_tie==1,]

# construct network
attr(friendship_edgelist, "n")=size

krack_friendship=network(x=friendship_edgelist, directed=T, 
                         vertex.attr=attribute_list)
```

Plot
```{r}
library(GGally)

# advice network
indeg_advice=degree(krack_advice, cmode="indegree")
ggnet2(krack_advice, node.color="LEVEL", node.size=indeg_advice,
       palette=c("1"="navy", "2"="blue", "3"="lightskyblue"), 
       edge.size=.5, arrow.size=3, arrow.gap=0.02, edge.color="grey80")+
  guides(size=FALSE)

# friendship network
indeg_friendship=degree(dat=krack_friendship, cmode="indegree")

ggnet2(krack_friendship, node.color="LEVEL", node.size= indeg_friendship,
       palette=c("1"="navy", "2"="blue", "3"="lightskyblue"), 
       edge.size=.5, arrow.size=3, arrow.gap=0.02, edge.color="grey80")+
  guides(size=FALSE)

```

## 2. Dyadic Processes

Dyad census: null, asymmetric, mutual

```{r}
dyadcensus_friendship=dyad.census(krack_friendship)
dyadcensus_friendship

recip_friendship=grecip(dat=krack_friendship, measure="dyadic.nonnull")
recip_friendship 

recip_advice=grecip(dat=krack_advice, measure="dyadic.nonnull")
recip_advice
```
- reciprocity rates similar for friendship and advice

Need to compare to baseline to really determing value/meaning
```{r}
cug_recip=cug.test(dat=krack_friendship, FUN=grecip,  mode=c("digraph"), 
                    cmode=c("edges"), reps=1000, 
                    FUN.args=list(measure="dyadic.nonnull"))
cug_recip

# observed
cug_recip$obs.stat

# generated null
summary(cug_recip$rep.stat)
```
- shows that observed reciprocity rate is higher
- network is influenced by reciprocity

```{r}
table(attributes$LEVEL)

# determine if ij pair in same level
same_level=outer(attributes$LEVEL, attributes$LEVEL, "==") 

dyad_mat_function=function(dat){#input is the network of interest

mat=as.matrix(dat) #getting matrix form of network
diag(mat)=NA #putting NA diagonal as we don’t want to consider self-ties

#Next, we do a little trick where we take the matrix and add it to its 
#transpose, yielding a matrix of 0s, 1s 
#and 2s. If it is null, the resulting value will be 0 
#(neither ij nor ji exists); if it is asymmetric there
#will be a 1 (as ij or ji exists but not both); and if it is mutual there
#will be a 2 (as there are ties from ij and ji).
dyad_mat=mat+t(mat) 

#Now we label for ease of interpretation:
dyad_mat[dyad_mat==0]="null" 
dyad_mat[dyad_mat==1]="asym"
dyad_mat[dyad_mat==2]="mut"
dyad_mat
}

friends_dyads=dyad_mat_function(krack_friendship)
friends_dyads[1:5, 1:5] 
```

```{r}
# dyad census
table(friends_dyads)/2

# for same level
table(friends_dyads[same_level==TRUE])/2

# not same level
table(friends_dyads[same_level==FALSE])/2
```
- shows how being on the same level affects relationship
- more mutual relationships when on same level

```{r}
advice_dyads=dyad_mat_function(krack_advice)
advice_dyads[1:5, 1:5]

table(friends_dyads, advice_dyads)/2
```

## 3. Triadic Processes

```{r}
# friendship
triads_friendship=triad.census(krack_friendship)

# advice
triads_advice=triad.census(krack_advice) 

# compare
triad_data= data.frame(advice=t(triads_advice), 
                       friendship=t(triads_friendship))
triad_data
```

```{r}
count_triads_function=function(dat, triad_names){#inputs are 
#the network and the name of the triad type
triads=triad.census(dat) #calculating the triad census
triads[1, triad_names] #grabbing counts of triad of interest
}

triads_to_count=c("030T")  

count_triads_function(krack_advice, triad_names=triads_to_count)
```

```{r}
cug_triads_advice=cug.test(krack_advice, FUN=count_triads_function,
                           mode=c("digraph"), 
                           cmode=c("dyad.census"), reps=1000, 
                           FUN.args=list(triad_names=triads_to_count)) 
cug_triads_advice
```
- num 030T ties is unusual

```{r}
cug_triads_friendship=cug.test(krack_friendship, FUN=count_triads_function, 
                               mode=c("digraph"), 
                               cmode=c( "dyad.census"), reps=1000, 
                               FUN.args=list(triad_names=triads_to_count))
cug_triads_friendship

```
- not unusual

```{r}
cug_triads_friendship$obs.stat
summary(cug_triads_friendship$rep.stat) 
```

```{r}
triads_to_count2=c("120U")

cug_triads_friendship2=cug.test(krack_friendship, FUN=count_triads_function, 
                                mode=c("digraph"),
                                cmode=c("dyad.census"), reps=1000, 
                                FUN.args=list(triad_names=triads_to_count2))
cug_triads_friendship2
```

```{r}
source("https://raw.githubusercontent.com/JeffreyAlanSmith/Integrated_Network_Science/master/R/CUGtest_alltriads.R")

CUG_test_table=CUGtest_alltriads(dat=krack_friendship, 
                                 mode="digraph",                      
                                 cmode="dyad.census", reps=1000)
CUG_test_table
```

```{r}
gtrans(dat=krack_advice, measure="weak") 
```

